Survival Analysis

### 2014-04-27 paper.Rnw
Tweak image and include algorithm steps

\usepackage[margin=1in]{geometry}
\usepackage{nicefrac}
\usepackage{tabularx}

@@

While many software packages can easily create survival estimates, computing $\hat S(t)$ can be relatively straightforward using tables. Let's examine sample data shown as Figure \ref{fig:ws}.

\begin{figure}[p]
<<ws, echo = FALSE, fig = TRUE, height = 3>>=
ggplot(ws, aes(y = id, yend = id, colour = event)) +
  ggtitle('Event Duration') +
  xlab('duration') +
  ylab('survival') +
  geom_segment(aes(
    x = 0,
    xend = duration,
    order = id)) +
  geom_point(aes(
    x = duration), size = 2)
@
\caption{thing}
\label{fig:ws}
\end{figure}

>>>>

\usepackage[margin=1in]{geometry}
\usepackage{nicefrac}
\usepackage{tabularx}
\usepackage{graphicx}

@@

While many software packages can easily create survival estimates, computing $\hat S(t)$ can be straightforward using tables. Let's examine sample data shown below.

<<ws, echo = FALSE, fig = TRUE, include = FALSE>>=
ggplot(ws, aes(y = id, yend = id, colour = event)) +
  theme(legend.position = "none") +
  ggtitle('Event Duration') +
  xlab('duration') +
  ylab('survival') +
  guides(fill = FALSE) +
  geom_segment(aes(
    x = 0,
    xend = duration,
    order = id)) +
  geom_point(aes(
    x = duration), size = 2)
@
\begin{figure}[h]
\begin{center}
\includegraphics[keepaspectratio=true, width=3in]{paper-ws.pdf}
\end{center}
\caption{Sample Survival Data}
\label{fig:ws}
\end{figure}

In this sample data, teal endpoints represent an event observation and red endpoints repreesnt a censored observation. We begin by creating a table with the following quantities: $n_i$, $d_i$, $c_i$, $\nicefrac{d_i}{n_i}$, $\nicefrac{n_i - d_i}{n_i}$, and $\hat S(t_i)$. Afterwards, we may perform the following algorithm:

\begin{enumerate}
  \item \emph{Calculate $n_i$.} $n_i$ is the total number of subjects that are at risk in this interval. If a vertical line is drawn at $t_i$, $n_i$ is the total number of duration lines the vertical line touches. This includes subjects that either have an event or are censored in this interval. Equivalently, you may also calculate $n_i$ by taking $n_{i - 1}$ and subtracting the number of events and censors in the previous duration.
  \item \emph{Calculate $d_i$.} $d_i$ is the number of teal duration endings at $t_i$.
  \item \emph{Calculate $c_i$.} $c_i$ is the number of red duration endings at $t_i$.
  \item \emph{Calculate $\nicefrac{d_i}{n_i}$ and $\nicefrac{n_i - d_i}{d_i}$} where the former represents the failure rate $f(t_i)$ and the latter represents the probability of not failing at that specific interval. Only the latter is used to calculate $\hat S(t_i)$.
  \item \emph{Calculate $S(t_i)$.} Recall that $S(t_i) = \nicefrac{n_i - d_i}{d_i} \cdot S(t_{i - 1})$. If $S(t_{i - 1})$ is not defined, use 1 indicating that everyone survived the imaginary duration.
  \item \emph{Repeat until $n_i$ is zero.}
\end{enumerate}
